Quantile Forecast Matching with a Bayesian Quantile Gaussian Process Model

### 2025-01-06 quantile_matching.tex
fix latex

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsmath package provides various useful equation environments.
\usepackage{amsmath}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

@@

@@

\begin{figure}
    \centering
    \includegraphics[scale=.5]{Images/fit_quantiles_example.png}
    \caption{Points representing 12 quantiles ($y$-axis) for given probability values ($x$-axis) estimated from a random sample of size 100 from a standard normal distribution. The quantile function (QF) of normal distribution (grey) was fit by selecting the mean and standard deviation parameters which minimize the least squares distance between the quantiles and the QF.}
    \label{fig:quant_match_example}
\end{figure}

@@

In this manuscript we introduce a novel model, the quantile Gaussian process (QGP), used specifically for QM. The context for using the QGP is that the available data is a set of quantiles estimated for a given set of probabilities.
The QGP relies on established theory underlying the relationship between the QF of a continuous distribution and sample quantiles \cite[]{parzen2004quantile, gilchrist2000statistical,hyndman1996sample,walker1968note,cramer1951mathematical}. The QGP can provide accurate asymptotic uncertainty quantification on quantiles, and it works well for QM for quantile defined distributions. 
In section \ref{sec:quant_func} the QF of a continuous random variable and sample quantiles are each defined, and important properties and asymptotic theory are reviewed. In section \ref{sec:qgp_model} the QGP model for QM is introduced. Section \ref{sec:simulation_analyses} contains simulation studies illustrating the QGP's ability to make parameter inference and match the distribution from which the quantiles where estimated. Section \ref{sec:qm_analysis} contains an application of QM using the QGP model for quantile forecasts of the 2023-24 United States Centers for Disease Control (CDC) flu forecasting competition, also known as FluSight. Section \ref{sec:qm_conclusion} concludes this manuscript with a short summary and ideas for further development in QM modeling.

@@

\begin{definition}
    \label{def:qf}
    For a cumulative distribution function $F: \mathbb{R} \rightarrow [0,1]$, the quantile function $Q: [0,1] \rightarrow \mathbb{R}$ is defined as

    \begin{equation*}
    \label{eq:qf}
        Q(p) = F^{-1}(p) := \emph{inf}\{x \in \mathbb{R} : F(x) \geq p\}, \quad p \in [0,1]
    \end{equation*}
    An alternative definition not in terms of the cumulative distribution function is a function $Q: [0,1] \rightarrow \mathbb{R} \cup \{-\infty, \infty\}$, which is nondecreasing and left-continuous. 
\end{definition}
% \jarad{the latter does not sound like a definition to me, but instead lists two properties.}

@@

Besides the CDF, PDF, and QF a  fourth infrequently mentioned defining function of a continuous random variable is the quantile density function (QDF). As the PDF is the derivative of the CDF, the QDF $q:[0,1] \rightarrow \mathbb{R}$ is the derivative of the QF defined as
\[
    q(p) = \frac{dQ(p)}{dp}.
\]
If $Q(p)$ is the inverse of a CDF $F(x)$, a calculus result shows that the reciprocal QDF $[q(p)]^{-1} = f(Q(p))$, or the PDF evaluated at the QF at $p$ \cite[]{perepolkin2023tenets, gilchrist2000statistical}. The QDF is important for establishing the CLTs in section \ref{sec:samp_quants} and for the QGP models introduced in section \ref{sec:qgp_model}.

@@

\begin{definition}
For $n$ independent observations $Y_1, ..., Y_n$ from a distribution with corresponding order statistics $Y_{(1)}, ..., Y_{(n)}$, the sample quantile function $\hat{Q}_n(p)$ may be defined as
    \label{def:sqf}
    \begin{equation}
        \label{eq:quant_df}
        \hat{Q}_n(p) = (1- \gamma)Y_{(j)} + \gamma Y_{(j + 1)}
    \end{equation}
    where 
    \[\frac{j - m}{n} \leq p < \frac{j - m + 1}{n}
    \]
    for some $m \in \mathbb{R}$ and $0 \leq \gamma \leq 1$ where $\gamma$ is some function of $j = \lfloor pn + m \rfloor$,  $g = pn + m - j$, and $\lfloor \cdot \rfloor$ is the floor function.
\end{definition}

@@

\begin{equation}
    \label{eq:qclt}
    \sqrt{n}(\hat{\boldsymbol{Q}}_n(\boldsymbol{p}) - \boldsymbol{Q}(\boldsymbol{p})) \overset{D}{\rightarrow} N_K(0, \boldsymbol{\mathcal{K}})
\end{equation}
where $\boldsymbol{\mathcal{K}}$ is a covariance matrix with the $ij^{th}$ entry $\kappa_{ij}$ where

@@

\begin{equation}
    \label{eq:qgp}
\hat{\boldsymbol{Q}}_n(\boldsymbol{p}) \sim N_K(\boldsymbol{Q}_{\theta}(\boldsymbol{p}), n^{-1} \boldsymbol{\mathcal{K}}_{\theta})
\end{equation}
Here $\theta$ is an unknown parameter to be estimated from the data. The covariance matrix $\boldsymbol{\mathcal{K}}_{\theta}$ has entries from the covariance function $\kappa_{\theta}(\cdot, \cdot)$ where 
\[
\kappa_{\theta}(p, p') = \frac{p\wedge p' - p p'}{f_{\theta}(Q_{\theta}(p)) f_{\theta}(Q_{\theta}(p'))}, \quad p, p' \in (0,1)
\]

@@

\begin{equation}
    \label{eq:norm_qgp}
\hat{\boldsymbol{Q}}_n(\boldsymbol{p}) \sim N_K \left(\mu + \sigma \boldsymbol{\Phi}^{-1}(\boldsymbol{p}), 
\frac{\sigma^2}{n} \boldsymbol{\Psi}}\right)
\end{equation}
Here $\boldsymbol{\Phi}^{-1}(\boldsymbol{p}) = (\Phi^{-1}(p_1), ..., \Phi^{-1}(p_K))$ is a known vector since $\boldsymbol{p}$ is given, and $\boldsymbol{\Psi}$ is a known $K\times K$ matrix with $ij^{th}$ entry

@@

Take for instance the data $\boldsymbol{X} = \left(\textbf{1}, \Phi^{-1}(\boldsymbol{p})\right)$ and the parameter vector $\beta = (\mu, \sigma)$. Assuming the sample size $n$ is unknown and setting $\sigma^2/n = \gamma^2$, model (\ref{eq:norm_qgp}) then becomes
\[
    \hat{\boldsymbol{Q}}_n(\boldsymbol{p}) \sim N_K\left(X\beta, 
\gamma^2 \boldsymbol{\Psi}}\right)
\]
In this model, one simplifying assumption may be to treat $\sigma$ and $\gamma^2$ separately and proceed to fit a standard linear regression model. All frequentist and Bayesian results of the linear regression model apply, including the existence of conditionally conjugate prior distributions. The positive constraint on $\sigma$ can also be dealt with without adding much complication (see \cite{gelman2013bayesian} pgs. 377-378). If for a certain problem it is important to make inference on the unknown $n$, or $\sigma$ and $\gamma^2$ cannot treated as separate, then one may estimate the two parameters by assigning appropriate prior distributions to $\sigma$ and $n$ and reverting back to (\ref{eq:norm_qgp}). In section \ref{sec:simulation_analyses}, we analyze a normal QGP in a simulation study where we assign independent prior distributions to $\sigma$ and $n$, and $\sigma$ is treated as both a regression coefficient and as part of the variance.

@@

Often $F_{\theta}$ will not be easily invertible, thus $Q_{\theta}$ will be a complicated function which can be evaluated only via numerical optimization which is often computationally expensive and/or inaccurate. A simple solution is to model the PIT transformed quantiles from the data rather than modeling the quantiles directly. Model (\ref{eq:qgp}) is then reformulated to be model (\ref{eq:pit_qgp}), where  $\Gamma_{ij} = p_i\wedge p_j - p_i p_j$ as in (\ref{eq:tqclt}).

@@

In section \ref{sec:simulation_analyses}, we analyze model (\ref{eq:pit_qgp}) where $F_{\theta}$ is set as a normal mixture distribution with $C = 4$ component normal distributions. Fitting this model proved faster and produced better results than trying to fit model (\ref{eq:qgp}) and evaluating $Q_{\theta}$ the QF of a normal mixture. In fact we found that in cases where $F_{\theta}$ was a simpler function, such as a normal or exponential distribution, fitting model (\ref{eq:pit_qgp}) was at least as fast and the results were at least as good as fitting (\ref{eq:qgp}). Thus when possible, we elected to fit (\ref{eq:pit_qgp}).

@@

The final method we include is the ORD model introduced by \cite{nirwan2020bayesian}. This model relies on the definition of sample quantiles being order statistics. In the ORD model, the joint distribution of a set of order statistics is the model likelihood. The likelihood of a set of order statistics from a continuous distribution is a function of both the CDF and PDF of a continuous distribution. ORD is the method most similar to QGP and it provides exact inference of quantile uncertainty whereas the QGP provides asymptotic uncertainty.

@@

\begin{equation}
    \label{eq:wass_dist_cdf}
    WD_p (F_X, F_Y) = \left( \int_{\mathbb{R}} |F_X(t) - F_Y(t)|^p dt\right)^{1/p}
\end{equation}

@@

\subsection{Parameter estimation and quantile matching for known distribution families}

Where the distribution family is known, one goal of QM may be to estimate and make inference on model parameters. In the first two studies, we analyzed parameter estimation and inference for normal and exponential family distributions with known parameters. We analyzed the QGP model from (\ref{eq:pit_qgp}), the IND model, and the ORD model. We note that model (\ref{eq:pit_qgp}) allows one to do inference for both the model parameter $\theta$ and for an unknown sample size $n$. The ORD model also allows for estimation of an unknown $n$. For the QGP and ORD models, we include modeling cases where $n$ is known and where $n$ is unknown and estimated. We denote the models where $n$ is known as QGPN and ORDN. We also compared the distance between the true model and the predictive distributions estimated via QGP, IND, ORD, SPL, and KDE. We also include a comparison of the QGP and ORD models where the true distribution is a quantile defined distribution.

@@

\begin{align*}
    \mu &\sim N(5,7^2) \\
    \sigma &\sim N(0, 6^2)\mathbbm{1}\{\sigma > 0\} \\
    n &\sim N(0,3000^2)\mathbbm{1}\{n > 0\}.
\end{align*}
For the IND model the prior on the independent parameter $\sigma_{\rho}\mathbbm{1}\{n > 0\}$ from model (\ref{eq:ind_quantile}) is $1/\sigma_{\rho} \sim N(0, 3000^2)$, which is similar to the prior on $n$ for the QGP and ORD models. For each fit, the HMC chain was run for 60,000 draws with the first 10,000 discarded as a burn-in.

Figure \ref{fig:normal_densities} shows examples of marginal posterior densities for $\mu$, $\sigma$, and unknown $n$ where the true $n \in \{50,\,150,\,500,\,1,000\}$ and $K \in \{7,\,15,\,23\}$. 
The parameter uncertainty of the QGP and ORD models are unsurprisingly similar in most cases, and where they differ it is hard to say that one is estimating the true parameter better than the other. The IND model on the other hand has tighter densities, and often the bulk of the posterior is far from the true parameter.
\begin{figure}[ht!]
\begin{subfigure}{.475\linewidth}
  \includegraphics[width=\linewidth]{Images/normal_mean_fit.png}
  % \caption{}
  % \label{MLEDdet}
\end{subfigure}\hfill % <-- "\hfill"
\begin{subfigure}{.52\linewidth}
  \includegraphics[width=\linewidth]{Images/normal_sd_fit.png}
  % \caption{}
  % \label{energydetPSK}
\end{subfigure}

\medskip % create some *vertical* separation between the graphs
\begin{subfigure}{.53\linewidth}
  \includegraphics[width=\linewidth]{Images/normal_n_fit.png}
  % \caption{}
  % \label{velcomp}
\end{subfigure}\hfill % <-- "\hfill"
\tiny
\caption{Density plots of posterior distribution samples for normal parameters by QM for QGP, ORD, and IND models. QM was done on estimated quantiles from a normal distribution with mean 4 and standard deviation 3.5. The posterior densities are for $\mu$ (top left), $\sigma$ (top right), and sample size $n$ (bottom). Plots are faceted by the sample size $n \in \{50, 150, 500, 1{,}000\}$ ($y$-axis) and number of quantiles $K \in \{7, 15, 23\}$ ($x$-axis). Vertical lines (black) show the value of the true parameter.}
\label{fig:normal_densities}
\end{figure}

The left side of figure \ref{fig:normal_cov_dists} shows the 90\% coverage of the posterior distribution credible intervals as $n$ increases for $\mu$, $\sigma$ and $n$ for five models. QGPN and ORDN models where $n$ is known are included along with QGP, ORD, and IND. 90\% credible intervals of the parameters were calculated by computing the $0.05^{th}$ and $0.95^{th}$ quantiles from the posterior distribution samples. The coverage percentage is calculated as the percentage of the 500 replicates which the true parameter value was within the 90\% credible interval.
As expected, the nominal coverage for the QGP and ORD models is better than that of the IND model, and for the two models where $n$ is known the coverage is better than where it is unknown, particularly for lower values of $n$ and $K$. The coverage for the ORD models appears to be slightly better than for the QGP models though not by much and not for every combination of $n$ and $K$. This is likely because the ORD model provides exact inference whereas the QGP model provides asymptotic inference.

\begin{figure}[hbt!]
\begin{subfigure}{.465\linewidth}
  \includegraphics[width=\linewidth]{Images/normal_coverage.png}
  % \caption{}
  % \label{MLEDdet}
\end{subfigure}\hfill % <-- "\hfill"
\begin{subfigure}{.524\linewidth}
  \includegraphics[width=\linewidth]{Images/normal_dists.png}
  % \caption{}
  % \label{energydetPSK}
\end{subfigure}
\caption{Posterior coverage (left) calculated as the percentage of times the true parameter fell within the modeled 90\% credible interval over the 500 replications. Coverage is faceted by the normal parameters $\mu$, $\sigma$, and $n$ with $K \in \{3, 7, 15, 23, 50\}$, and by increasing sample size ($x$-axis). The five models QGP, ORD, QGPN, ORDN, and IND are colored as shown the legend. The horizontal line (black) is at the nominal 90\% level. Only QGP and ORD appear for the parameter $n$ as they are the only two which estimate an unknown $n$.
Distance between the true distribution and the estimated QM predictive distribution (right) averaged over the 500 replications. Distances include the UWD1, TV, and KLD for $K \in \{3, 7, 15, 23, 50\}$, and by increasing sample size ($x$-axis).}
\label{fig:normal_cov_dists}
\end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[scale=.6]{Images/normal_coverage.png}
%     \caption{Percentage of times of the 1,000 simulation replicates where the true sample fell within the modeled 90\% credible interval. The coverage is faceted by the normal parameters $\mu$, $\sigma$, and $n$ with $K \in \{3, 7, 15, 23, 50\}$ on the $y$-axis, and by increasing sample size. The five models QGP, ORD, QGPN, ORDN, and IND are colored as shown the legend. The black vertical line is at the nominal 90\% level. Only QGP and ORD appear for the parameter $n$ as they are the only two which estimate an unknown $n$.}
%     \label{fig:norm_cover}
% \end{figure}

To analyze how well QM methods produced predictive distributions which approximate the true distribution $N(4, 3.5^2)$, we measured the distance between the estimated predictive distributions of QGP, ORD, and IND fits to the true distribution. 
We also measured the distances of the SPL and KDE QM fits to the true distribution. The distances were measured via the UWD1, TV, and KLD metrics. 
The UWD1 for each of QGP, ORD, QGPN, ORDN, and IND were calculated as follows. For $\{\theta_m\}^M$ being $M$ draws from the posterior distribution of $\theta$ where $\theta_m = (\mu_m, \sigma_m)$, the QM posterior predictive distribution was simulated by repeatedly sampling a $\theta_m^*$, then sampling $X_m$ from a distribution with CDF $F_{\theta_m^*}$. Repeating this for $M = 50,000$ times gives a QM posterior predictive sample. For the CDF $F_{\theta}$ where $\theta$ is the true parameter, $\xi_m = F_{\theta}(X_m)$ was calculated and the empirical CDF $\hat{F}_{\xi}$ was calculated from the sample $\{\xi_m\}^M$. $F_{\xi}$ in (\ref{eq:pit_wass_dist}) was then replaced by $\hat{F}_{\xi}$ to calculate UWD1. For SPL and KDE $F_{\xi}(x)$ was replaced by the respective estimated CDFs. The \texttt{integrate} function in \texttt{R} was used for calculating the UWD1 in (\ref{eq:pit_wass_dist}).

To estimate TV and KLD, rather than sampling from the posterior distribution, the marginal means for the parameters were calculated from the posterior distribution samples. These marginal means  are $\Bar{\theta}_M = (\Bar{\mu}_M, \Bar{\sigma}_M)$ where $\Bar{\mu}_M$ and $\Bar{\sigma}_M$. To calculate TV, $f$ in (\ref{eq:tot_var}) is replaced by $f_{\Bar{\theta}_M}$ and $g$ is replaced by $f_{\theta}$. The TV calculation was done using the \texttt{integrate} function in \texttt{R}. The KLD was estimated also using $f_{\Bar{\theta}_M}$ and $f_{\theta}$, but was done via importance sample. That is (\ref{eq:kld}) was estimated by

@@

This study shows the ability of the QGP QM model to estimate and perform inference on model parameters where the true distribution family is known. It also shows the QGP's ability to produce a posterior predictive distribution which closely matches a true distribution, according to several metrics, relative to other QM methods. Between parameter inference and predicting the true distribution, QGP is superior to all methods we compared it with except for the ORD model which has similar results to the QGP. A similar simulation study where the data is simulated from an exponential family distribution is in appendix \ref{app:extra_sim}. The results of that study are similar to those of the normal distribution family study.

>>>>

%% The amssymb package provides various useful mathematical symbols
% \usepackage{amssymb}
%% The amsmath package provides various useful equation environments.
% \usepackage{amsmath}
\usepackage{amsmath,amsthm,amssymb,scrextend,mathtools}
\usepackage{hyperref}
\usepackage{newtxmath} %for making indicator function
% \RequirePackage[colorlinks,citecolor=blue,urlcolor=blue,backref=page,backref=page]{hyperref}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

@@

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\usepackage{natbib}

@@

\begin{figure}[hbt!]
    \centering
    \includegraphics[scale=.5]{Images/fit_quantiles_example.png}
    \caption{Points representing 12 quantiles ($y$-axis) for given probability values ($x$-axis) estimated from a random sample of size 100 from a standard normal distribution. The quantile function (QF) of normal distribution (grey) was fit by selecting the mean and standard deviation parameters which minimize the least squares distance between the quantiles and the QF.}
    \label{fig:quant_match_example}
\end{figure}

@@

In this manuscript we introduce a novel model, the quantile Gaussian process (QGP), used specifically for QM. The context for using the QGP is that the available data is a set of quantiles estimated for a given set of probabilities.
The QGP relies on established theory underlying the relationship between the QF of a continuous distribution and sample quantiles \cite[]{parzen2004quantile, gilchrist2000statistical,hyndman1996sample,walker1968note,cramer1951mathematical}. The QGP can provide accurate asymptotic inference of uncertainty of quantiles, and it works well for QM for quantile defined distributions. 
In section \ref{sec:quant_func} the QF of a continuous random variable and sample quantiles are each defined, and important properties and asymptotic theory are reviewed. In section \ref{sec:qgp_model} the QGP model for QM is introduced. Section \ref{sec:simulation_analyses} contains simulation studies illustrating the QGP's ability to make parameter inference and match the distribution from which the quantiles where estimated. Section \ref{sec:qm_analysis} contains an application of QM using the QGP model for quantile forecasts of the 2023-24 United States Centers for Disease Control (CDC) flu forecasting competition, also known as FluSight. Section \ref{sec:qm_conclusion} concludes this manuscript with a short summary and ideas for further development in QM modeling.

@@

\begin{definition}
    \label{def:qf}
    For a cumulative distribution function $F: \vmathbb{R} \rightarrow [0,1]$, the quantile function $Q: [0,1] \rightarrow \vmathbb{R}$ is defined as

    \begin{equation*}
    \label{eq:qf}
        Q(p) = F^{-1}(p) := \emph{inf}\{x \in \vmathbb{R} : F(x) \geq p\}, \quad p \in [0,1]
    \end{equation*}
    An alternative definition not in terms of the cumulative distribution function is a function $Q: [0,1] \rightarrow \vmathbb{R} \cup \{-\infty, \infty\}$, which is nondecreasing and left-continuous. 
\end{definition}
% \jarad{the latter does not sound like a definition to me, but instead lists two properties.}

@@

Besides the CDF, PDF, and QF a  fourth infrequently mentioned defining function of a continuous random variable is the quantile density function (QDF). As the PDF is the derivative of the CDF, the QDF $q:[0,1] \rightarrow \vmathbb{R}$ is the derivative of the QF defined as
\[
    q(p) = \frac{dQ(p)}{dp}.
\]
If $Q(p)$ is the inverse of a CDF $F(x)$, a calculus result shows that the reciprocal QDF $[q(p)]^{-1} = f(Q(p))$, or the PDF evaluated at the QF at $p$ \cite[]{perepolkin2023tenets, gilchrist2000statistical}. The QDF is important for establishing the CLTs in section \ref{sec:samp_quants} and for the QGP models introduced in section \ref{sec:qgp_model}.

@@

\begin{definition}
For $n$ independent observations $Y_1, ..., Y_n$ from a distribution with corresponding order statistics $Y_{(1)}, ..., Y_{(n)}$, the sample quantile function $\hat{Q}_n(p)$ may be defined as
    \label{def:sqf}
    \begin{equation}
        \label{eq:quant_df}
        \hat{Q}_n(p) = (1- \gamma)Y_{(j)} + \gamma Y_{(j + 1)}
    \end{equation}
    where 
    \[\frac{j - m}{n} \leq p < \frac{j - m + 1}{n}
    \]
    for some $m \in \vmathbb{R}$ and $0 \leq \gamma \leq 1$ where $\gamma$ is some function of $j = \lfloor pn + m \rfloor$,  $g = pn + m - j$, and $\lfloor \cdot \rfloor$ is the floor function.
\end{definition}

@@

\begin{equation}
    \label{eq:qclt}
    \sqrt{n}(\hat{\boldsymbol{Q}}_n(\boldsymbol{p}) - \boldsymbol{Q}(\boldsymbol{p})) \overset{D}{\rightarrow} N_K(0, \boldsymbol{\mathcal{K}})
\end{equation}
where $\boldsymbol{\mathcal{K}}$ is a $K\times K$ with the $ij^{th}$ entry $\kappa_{ij}$ where

@@

\begin{equation}
    \label{eq:qgp}
\hat{\boldsymbol{Q}}_n(\boldsymbol{p}) \sim N_K(\boldsymbol{Q}_{\theta}(\boldsymbol{p}), n^{-1} \boldsymbol{\mathcal{K}}_{\theta})
\end{equation}
Here $\theta$ is an unknown parameter to be estimated from the data. The covariance matrix $\boldsymbol{\mathcal{K}}_{\theta}$ has entries from the covariance function $\kappa_{\theta}(\cdot, \cdot)$ where 
\[
\kappa_{\theta}(p, p') = \frac{p\wedge p' - p p'}{f_{\theta}(Q_{\theta}(p)) f_{\theta}(Q_{\theta}(p'))}, \quad p, p' \in (0,1)
\]
Here the operator $a\wedge b$ indicates the minimum value between $a$ and $b$.

@@

\begin{equation}
    \label{eq:norm_qgp}
\hat{\boldsymbol{Q}}_n(\boldsymbol{p}) \sim N_K \left( \mu + \sigma \boldsymbol{\Phi}^{-1}(\boldsymbol{p}),\frac{\sigma^2}{n} \boldsymbol{\Psi} \right)
\end{equation}
Here $\boldsymbol{\Phi}^{-1}(\boldsymbol{p}) = (\Phi^{-1}(p_1), ..., \Phi^{-1}(p_K))$ is a known vector since $\boldsymbol{p}$ is given, and $\boldsymbol{\Psi}$ is a known $K\times K$ matrix with $ij^{th}$ entry

@@

Take for instance the data $\boldsymbol{X} = \left(\textbf{1}, \Phi^{-1}(\boldsymbol{p})\right)$ and the parameter vector $\beta = (\mu, \sigma)$. Assuming the sample size $n$ is unknown and setting $\sigma^2/n = \gamma^2$, model (\ref{eq:norm_qgp}) then becomes
\[
    \hat{\boldsymbol{Q}}_n(\boldsymbol{p}) \sim N_K\left(X\beta, 
\gamma^2 \boldsymbol{\Psi}\right)
\]
In this model, one simplifying assumption may be to treat $\sigma$ and $\gamma^2$ separately and proceed to fit a standard linear regression model. All frequentist and Bayesian results of the linear regression model apply, including the existence of conditionally conjugate prior distributions. The positive constraint on $\sigma$ can also be dealt with without adding much complication (see \cite{gelman2013bayesian} pgs. 377-378). If for a certain problem it is important to make inference on the unknown $n$, or $\sigma$ and $\gamma^2$ cannot treated as separate, then one may estimate the two parameters by assigning appropriate prior distributions to $\sigma$ and $n$ and reverting back to (\ref{eq:norm_qgp}). In section \ref{sec:simulation_analyses}, we analyze a normal QGP in a simulation study where we assign independent prior distributions to $\sigma$ and $n$, and $\sigma$ is treated as both a regression coefficient and as part of the variance.

@@

Often $F_{\theta}$ will not be easily invertible, thus $Q_{\theta}$ will be a complicated function which can be evaluated only via numerical optimization which is often computationally expensive and/or inaccurate. A simple solution is to model the PIT transformed quantiles from the data rather than modeling the quantiles directly. The QGP model (\ref{eq:qgp}) is then reformulated to be model (\ref{eq:pit_qgp}), where  $\Gamma_{ij} = p_i\wedge p_j - p_i p_j$ as in (\ref{eq:tqclt}).

@@

In section \ref{sec:simulation_analyses}, we analyze PIT transformed QGP model (\ref{eq:pit_qgp}) where $F_{\theta}$ is set as a normal mixture distribution with $C = 4$ component normal distributions. Fitting this model proved faster and produced better results than trying to fit the QGP model (\ref{eq:qgp}) and evaluating $Q_{\theta}$ the QF of a normal mixture. In fact we found that in cases where $F_{\theta}$ was a simpler function, such as a normal or exponential distribution, fitting model (\ref{eq:pit_qgp}) was at least as fast and the results were at least as good as fitting (\ref{eq:qgp}). Thus when possible, we elected to fit (\ref{eq:pit_qgp}).

@@

The final method we include is the ORD model introduced by \cite{nirwan2020bayesian}. This model relies on the definition of sample quantiles being order statistics. In the ORD model, the joint distribution of a set of order statistics is the model likelihood. The likelihood of a set of order statistics from a continuous distribution is a function of both the CDF and PDF of a continuous distribution. ORD is the method most similar to QGP and it provides exact inference of quantile uncertainty whereas the QGP provides asymptotic inference of uncertainty.

@@

\begin{equation}
    \label{eq:wass_dist_cdf}
    WD_p (F_X, F_Y) = \left( \int_{\vmathbb{R}} |F_X(t) - F_Y(t)|^p dt\right)^{1/p}
\end{equation}

@@

\subsection{Parameter estimation and quantile matching for known distribution \\families}

Where the distribution family is known, one goal of QM may be to estimate and make inference on model parameters. In the first two studies, we analyzed parameter estimation and inference for normal and exponential family distributions with known parameters. We analyzed the QGP model from (\ref{eq:pit_qgp}), the IND model, and the ORD model. We note that model (\ref{eq:pit_qgp}) allows one to do inference for both the model parameter $\theta$ and for an unknown sample size $n$. The ORD model also allows for estimation of an unknown $n$. For the QGP and ORD models, we include modeling cases where $n$ is known and where $n$ is unknown and estimated. We denote the models where $n$ is known as QGP-n and ORD-n. We also compared the distance between the true model and the predictive distributions estimated via QGP, IND, ORD, SPL, and KDE. We also include a comparison of the QGP and ORD models where the true distribution is a quantile defined distribution.

@@

\begin{align*}
    \mu &\sim N(5,7^2) \\
    \sigma &\sim N(0, 6^2) \vmathbb{1}\{\sigma > 0\}\\
    n &\sim N(0,3000^2)\vmathbb{1}\{n > 0\}.
\end{align*}
For the IND model the prior on the independent parameter $\sigma_{\rho}\vmathbb{1}\{n > 0\}$ from model (\ref{eq:ind_quantile}) is $1/\sigma_{\rho} \sim N(0, 3000^2)$, which is similar to the prior on $n$ for the QGP and ORD models. For each fit, the HMC chain was run for 60,000 draws with the first 10,000 discarded as a burn-in.

Figure \ref{fig:normal_densities} shows examples of marginal posterior densities for $\mu$, $\sigma$, and unknown $n$ where the true $n \in \{50,\,150,\,500,\,1,000\}$ and $K \in \{7,\,15,\,23\}$. 
The parameter uncertainty of the QGP and ORD models are unsurprisingly similar in most cases, and where they differ it is hard to say that one is estimating the true parameter better than the other. The IND model on the other hand has tighter densities, and often the bulk of the posterior is far from the true parameter.
\begin{figure}[hbt!]
% \begin{subfigure}{.475\linewidth}
  \includegraphics[width=.4\linewidth]{Images/normal_mean_fit.png}
  % \caption{}
  % \label{MLEDdet}
% \end{subfigure}\hfill % <-- "\hfill"
% \begin{subfigure}{.52\linewidth}
  \includegraphics[width=.4\linewidth]{Images/normal_sd_fit.png}
  % \caption{}
  % \label{energydetPSK}
% \end{subfigure}

\medskip % create some *vertical* separation between the graphs
%\begin{subfigure}{.53\linewidth}
  \includegraphics[width=.4\linewidth]{Images/normal_n_fit.png}
  % \caption{}
  % \label{velcomp}
%\end{subfigure}\hfill % <-- "\hfill"
\tiny
\caption{Density plots of posterior distribution samples for normal parameters by QM for QGP, ORD, and IND models. QM was done on estimated quantiles from a normal distribution with mean 4 and standard deviation 3.5. The posterior densities are for $\mu$ (top left), $\sigma$ (top right), and sample size $n$ (bottom). Plots are faceted by the sample size $n \in \{50, 150, 500, 1{,}000\}$ ($y$-axis) and number of quantiles $K \in \{7, 15, 23\}$ ($x$-axis). Vertical lines (black) show the value of the true parameter.}
\label{fig:normal_densities}
\end{figure}

The left side of figure \ref{fig:normal_cov_dists} shows the 90\% coverage of the posterior distribution credible intervals as $n$ increases for $\mu$, $\sigma$ and $n$ for five models. QGP-n and ORD-n models where $n$ is known are included along with QGP, ORD, and IND. 90\% credible intervals of the parameters were calculated by computing the $0.05^{th}$ and $0.95^{th}$ quantiles from the posterior distribution samples. The coverage percentage is calculated as the percentage of the 500 replicates which the true parameter value was within the 90\% credible interval.
As expected, the nominal coverage for the QGP and ORD models is better than that of the IND model, and for the two models where $n$ is known the coverage is better than where it is unknown, particularly for lower values of $n$ and $K$. The coverage for the ORD models appears to be slightly better than for the QGP models though not by much and not for every combination of $n$ and $K$. This is likely because the ORD model provides exact inference whereas the QGP model provides asymptotic inference.

\begin{figure}[hbt!]
%\begin{subfigure}{.465\linewidth}
  \includegraphics[width=.4\linewidth]{Images/normal_coverage.png}
  % \caption{}
  % \label{MLEDdet}
%\end{subfigure}\hfill % <-- "\hfill"
%\begin{subfigure}{.524\linewidth}
  \includegraphics[width=.4\linewidth]{Images/normal_dists.png}
  % \caption{}
  % \label{energydetPSK}
%\end{subfigure}
\caption{Posterior coverage (left) calculated as the percentage of times the true parameter fell within the modeled 90\% credible interval over the 500 replications. Coverage is faceted by the normal parameters $\mu$, $\sigma$, and $n$ with $K \in \{3, 7, 15, 23, 50\}$, and by increasing sample size ($x$-axis). The five models QGP, ORD, QGP-n, ORD-n, and IND are colored as shown the legend. The horizontal line (black) is at the nominal 90\% level. Only QGP and ORD appear for the parameter $n$ as they are the only two which estimate an unknown $n$.
Distance between the true distribution and the estimated QM predictive distribution (right) averaged over the 500 replications. Distances include the UWD1, TV, and KLD for $K \in \{3, 7, 15, 23, 50\}$, and by increasing sample size ($x$-axis).}
\label{fig:normal_cov_dists}
\end{figure}

To analyze how well QM methods produced predictive distributions which approximate the true distribution $N(4, 3.5^2)$, we measured the distance between the estimated predictive distributions of QGP, ORD, and IND fits to the true distribution. 
We also measured the distances of the SPL and KDE QM fits to the true distribution. The distances were measured via the UWD1, TV, and KLD metrics. 
The UWD1 for each of QGP, ORD, QGP-n, ORD-n, and IND were calculated as follows. For $\{\theta_m\}^M$ being $M$ draws from the posterior distribution of $\theta$ where $\theta_m = (\mu_m, \sigma_m)$, the QM posterior predictive distribution was simulated by repeatedly sampling a $\theta_m^*$, then sampling $X_m$ from a distribution with CDF $F_{\theta_m^*}$. Repeating this for $M = 50,000$ times gives a QM posterior predictive sample. For the CDF $F_{\theta}$ where $\theta$ is the true parameter, $\xi_m = F_{\theta}(X_m)$ was calculated and the empirical CDF $\hat{F}_{\xi}$ was calculated from the sample $\{\xi_m\}^M$. $F_{\xi}$ in (\ref{eq:pit_wass_dist}) was then replaced by $\hat{F}_{\xi}$ to calculate UWD1. For SPL and KDE $F_{\xi}(x)$ was replaced by the respective estimated CDFs. The \texttt{integrate} function in \texttt{R} was used for calculating the UWD1 in (\ref{eq:pit_wass_dist}).

To estimate TV and KLD, rather than sampling from the posterior distribution, the marginal means for the parameters were calculated from the posterior distribution samples. These marginal means  are $\Bar{\theta}_M = (\Bar{\mu}_M, \Bar{\sigma}_M)$ where $\Bar{\mu}_M$ and $\Bar{\sigma}_M$. To calculate TV, $f$ in (\ref{eq:tot_var}) is replaced by $f_{\Bar{\theta}_M}$ and $g$ is replaced by $f_{\theta}$. The TV calculation was done using the \texttt{integrate} function in \texttt{R}. The KLD was estimated also using $f_{\Bar{\theta}_M}$ and $f_{\theta}$, but was done via Monte Carlo sampling. That is (\ref{eq:kld}) was estimated by

@@

This study shows the ability of the QGP QM model to estimate and perform inference on model parameters where the true distribution family is known. It also shows the QGP's ability to produce a posterior predictive distribution which closely matches a true distribution, according to several metrics, relative to other QM methods. Between parameter inference and predicting the true distribution, QGP is superior to all methods we compared it with except for the ORD model which has similar results to the QGP. A similar simulation study where the data is simulated from an exponential family distribution is in appendix 
% \ref{app:extra_sim}. 
The results of that study are similar to those of the normal distribution family study.

When fitting the Bayesian models both in the normal and in the exponential setting, the IND model tended to fit the fastest, followed by the ORD models, and finally the QGP models. The time to fit however was very fast, usually no more than seven seconds except for the case where $K = 50$ quantiles, in which case the QGP model where $n$ was unknown often took around 20 seconds to fit. However, with $K= 50$ the QGP-n model continued to fit very rapidly, fitting in under seven seconds.
